---
title       : Developing Data Products - Forecast of Heathcare vs GDP
subtitle    : 
author      : GM
job         : 
framework   : shower        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Forecasting escalating Healthcare costs vs GDP 

1. BLS Medical Expenditure Panel Survey (MEPS) account data from 2000-2010
2. GDP and topline healthcare spending fit to a time series
3. The app forecasts into the future both the GDP and Healthcare costs
4. Forecasts are plotted with the 95% confidence interval

--- .class #id 

## Results

1. Healthcare costs are forecast to expand at the current rate
2. Surprisinging GDP did not increase, instead the 95% interval widened
3. Subsequent slides show the R calculations
4. Check the Shiny app to follow along

--- .class #id 

## 50 year forecast server code run output

```{r, echo=FALSE}
## Implement caching
getDat <- function() {
  
  library(openxlsx)
  library(forecast)
  setwd('C:/Users/gautam.Mom81/Documents/courseraDataScienceSpecialization/developingDataProducts/project')
  
  dfile <- 'data/MEPS.xlsx'
  dat <-  openxlsx::read.xlsx(dfile, startRow=2, sheet=3)
  
  # add the GDP data
  g <- openxlsx::read.xlsx(dfile, sheet = 4, rows=17:20)
  g <- g[1,]
  colnames(g) <- colnames(dat)
  
  dat <- rbind(dat,g)
  categories <- dat[[1]]
  
  dat <-  dat[,-1]
  dat <- t(dat)  # transpose the data frame and remove the first column
  colnames(dat) <- categories
  
  dat <- data.frame(dat)
  dat$Year <- as.numeric(rownames(dat))
  colnames(dat)[36] <- 'GDP'
  datRet <- subset(dat, select=c('Year', 'GDP', 'Health'))
  datRetOuter <<-datRet
  datRet
}

getForecast <- function(numYears=40) {
  
  dat <- getDat()

  tsHlth <-  ts(dat$Health, start = 2000)
  tsGdp <-  ts(dat$GDP, start = 2000)
 
  fcastHlth <- data.frame(forecast(tsHlth,  h = numYears, level=95, allow.multiplicative.trend=TRUE))
  fcastGdp <- data.frame(forecast(tsGdp,  h = numYears, level=95, allow.multiplicative.trend=TRUE))
  
  f <- data.frame(Year = as.numeric(rownames(fcastHlth)),
                 Health = fcastHlth$Point.Forecast,
                 HealthLo = fcastHlth$Lo.95,
                 HealthHi = fcastHlth$Hi.95,
                 GDP = fcastGdp$Point.Forecast,
                 GDPLo = fcastGdp$Lo.95,
                 GDPHi = fcastGdp$Hi.95
                 )
  
}

df <- getForecast(50)
df2 <- subset(df,select=c('Year', 'GDP', 'Health'))
head(df2)
```


---

## Forecasting Code snippet

```{r, eval = FALSE}

  tsHlth <-  ts(dat$Health, start = 2000)
  tsGdp <-  ts(dat$GDP, start = 2000)
 
  fcastHlth <- data.frame(forecast(tsHlth,  h = numYears, level=95, allow.multiplicative.trend=TRUE))
  fcastGdp <- data.frame(forecast(tsGdp,  h = numYears, level=95, allow.multiplicative.trend=TRUE))

```

---


## Conclusion

Healthcare costs increase is unsustainable !!

